\propertitle

### 2019-06-10 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-05-17T19:34:12+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

What correlations are important for the description of the multi-neuronal
activity in a specific brain area? How does such activity change when
external stimuli or experimental conditions change? Does such activity
range over all its theoretically possible values, or only over a
submanifold thereof?

Answering this kind of questions always engages an element of uncertainty.
Our answers therefore involve experimental data, such as neuronal
recordings from specific brain areas, and also probabilistic assumptions --
probabilistic \enquote{models} -- about biological conditions and
mechanisms that cannot be experimentally ascertained.

Despite remarkable advances in recording technologies, the best
experimental measurements of neuronal activity can still only record a very
small sample of neurons compared to the number that constitutes a brain
region. Many probabilistic models focus on such samples only, somehow
neglecting, in their assumptions, that the recorded neurons are a sample
from a larger network. This kind of isolation assumptions sometimes escape
attention, being subtly hidden in the mathematics. Some probabilistic
models try to take also unrecorded neurons into account, but become very
complex in doing so.

In the present work we give an answer to this question: What kind of
inferences about the total activity of a large neuronal network can we make
by recording a very small sample thereof? Our answer comes from the
simple combination of basic sampling probability relations and the maximum-entropy
method.\mynote{shall we give a two-sentence summary of the main idea here?}

The maximum-entropy method\mynote{refs} has been used for a variety of
inferences about neuronal activity from various brain areas and about other
phenomena of importance to the neurosciences, for example gene regulation
and protein sequencing\mynote{what's this exactly? refs}. This method is
often used to test whether some statistics of the data, for example
second-order time correlatins, is sufficient for making other inferences
about the data. It can be considered as an approximation of probabilistic
models based on various assumptions of inferential
sufficiency\mynote{refs} -- for example of empirical moments or of long-run
frequencies.

@@

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-10T16:14:24+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

What correlations are important for the description of the multi-neuronal
activity in a specific brain area? How does such activity change when
external stimuli or experimental conditions change? Does such activity
range over all its mathematically possible values, or only over a
subset thereof?

Answering this kind of questions always engages an element of uncertainty.
Our answers therefore involve both experimental data, such as neuronal
recordings from specific brain areas, and probabilistic assumptions --
probabilistic \enquote{models} -- about biological conditions and
mechanisms that cannot be experimentally ascertained.

Despite remarkable advances in recording technologies, the best
experimental measurements of neuronal activity can still only record a very
small sample of neurons compared to the numbers that constitute one
functionally distinguished brain region. Many probabilistic models focus on
such samples only, somehow neglecting, in their assumptions, that the
recorded neurons are a sample from a larger network. This kind of isolation
assumptions sometimes escape attention, being subtly hidden in the
mathematics. Some probabilistic models try to take also unrecorded neurons
into account, but become very complex in doing so.

In the present work we give an answer to this question: What kind of
inferences about the \emph{total activity} of a large neuronal network can
we make by recording a very small sample thereof? The answer we'll give
comes from the simple combination of basic sampling probability relations
and the maximum-entropy method.\mynote{shall we give a two-sentence summary
  of the main idea here?}

The maximum-entropy method\mynote{refs} has been used for a variety of
inferences about the neuronal activity of various brain areas and about
other phenomena of importance to the neurosciences, for example gene
regulation and protein sequencing\mynote{what's this exactly? refs}. This
method is often used to test whether some statistics of the data, for
example second-order time correlatins, is sufficient for making other
inferences about the data. It can be considered as an approximation of
probabilistic models based on various assumptions of inferential
sufficiency\mynote{refs} -- for example of empirical moments or of long-run
frequencies.

@@

The maximum-relative-entropy distribution is, in the first two cases, an
approximation of the most probable frequency distribution $\yF$ or $\ynu$;
in the third case, an approximation of the probability distribution for
$\yA'$.

\section{Assumptions and limitations}
\label{sec:assumptions}

Main assumptions behind this belief distribution:

We are approximating our state of knowledge with a finitely exchangeable
one. In turn, this is numerically approximated by an infinitely
exchangeable one for $T$ large. But we don't really have an exchangeable
belief: our \dob\ that the activity at the next time bin will
differ from the activity at the present one is roughly proportional to the
difference in the two subsequent activities.

The formulae say that we have equal beliefs about the underlying network
states having the same activity. This isn't really our belief, for we
believe there are interactions between the neurons and subnetworks thereof.
